---
title: "Significant Earthquakes 1900-2020"
author: "Oscar Gauna"
date: "September 24, 2020"
output: ioslides_presentation
---
<style>
p.caption {
  font-size: 0.6em;
}
</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Summary

The Significant Earthquake Database contains information on destructive earthquakes from 2150 B.C. to the present that meet at least one of the following criteria: Moderate damage (approximately $1 million or more), 10 or more deaths, Magnitude 7.5 or greater, Modified Mercalli Intensity X or greater, or the earthquake generated a tsunami. In this report I present the data between 1900 and 2020.

## Significant Earthquakes 1900-2020
##### Hover over the dots for location info or click the region in the legend to add/remove dots.
```{r plot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Data source: National Geophysical Data Center / World Data Service (NGDC/WDS): Significant Earthquake Database. National Geophysical Data Center, NOAA. doi:10.7289/V5TD9V7K'}
library(dplyr)
library(plotly)

data <- read.csv("Worldwide-Earthquake-database.csv")
region_code <- c(10,15,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160,170)
region_name <- c("Central, Western and S. Africa","Northern Africa","Antarctica",
    "East Asia","Central Asia and Caucasus","Kamchatka and Kuril Islands",
    "S. and SE. Asia and Indian Ocean","Atlantic Ocean","Bering Sea","Caribbean",
    "Central America","Eastern Europe","Northern and Western Europe","Southern Europe",
    "Middle East","North America and Hawaii","South America","Central and South Pacific")

regions <- data.frame(code=region_code, name=region_name)

data_small <- data %>%
    select(YEAR, REGION_CODE, LOCATION_NAME, EQ_PRIMARY) %>%
        filter(YEAR > 1899 & EQ_PRIMARY != "NA" & REGION_CODE != "NA") %>%
            mutate(REGION_CODE=factor(REGION_CODE))

new <- data_small
new[] <- lapply(data_small, function(x) regions$name[match(x, regions$code)])

data_small$REGION_CODE <- new$REGION_CODE
data_small$hover <- with(data_small, paste(LOCATION_NAME, '<br>', "YEAR:", YEAR, '<br>', "MAGNITUDE:", EQ_PRIMARY))

title <- list(text="SIGNIFICANT EARTHQUAKES 1900-2020")
legend <- list(title=list(text="REGION"))
y <- list(title = "MAGNITUDE")

fig <- plot_ly(data_small, x=~YEAR, y=~EQ_PRIMARY, type="scatter", mode="markers", color=~REGION_CODE, text=~hover)
fig <- fig %>% layout(title=title, yaxis=y, legend=legend)
fig
```

## Appendix (Code)
```{css, echo = FALSE}
.tiny {
  font-size: 8px;
  line-height: 1.5;
}
```
```{r class.source="tiny", echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
library(dplyr)
library(plotly)

#Read source data and define region codes
data <- read.csv("Worldwide-Earthquake-database.csv")
region_code <- c(10,15,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160,170)
region_name <- c("Central, Western and S. Africa","Northern Africa","Antarctica",
    "East Asia","Central Asia and Caucasus","Kamchatka and Kuril Islands",
    "S. and SE. Asia and Indian Ocean","Atlantic Ocean","Bering Sea","Caribbean",
    "Central America","Eastern Europe","Northern and Western Europe","Southern Europe",
    "Middle East","North America and Hawaii","South America","Central and South Pacific")

regions <- data.frame(code=region_code, name=region_name)

#Subset the fields of interest
data_small <- data %>%
    select(YEAR, REGION_CODE, LOCATION_NAME, EQ_PRIMARY) %>%
        filter(YEAR > 1899 & EQ_PRIMARY != "NA" & REGION_CODE != "NA") %>%
            mutate(REGION_CODE=factor(REGION_CODE))

#Translate region codes to names
new <- data_small
new[] <- lapply(data_small, function(x) regions$name[match(x, regions$code)])

data_small$REGION_CODE <- new$REGION_CODE

#Create hovering text
data_small$hover <- with(data_small, paste(LOCATION_NAME, '<br>', "YEAR:", YEAR, '<br>', "MAGNITUDE:", EQ_PRIMARY))

#Create plot
title <- list(text="SIGNIFICANT EARTHQUAKES 1900-2020")
legend <- list(title=list(text="REGION"))
y <- list(title = "MAGNITUDE")

fig <- plot_ly(data_small, x=~YEAR, y=~EQ_PRIMARY, type="scatter", mode="markers", color=~REGION_CODE, text=~hover)
fig <- fig %>% layout(title=title, yaxis=y, legend=legend)
fig
```